---
title: "Homework Assignment 1"
subtitle: "Map Making practice"
author: "Amapola Garcia"
date: 2026-10-6
execute: 
  eval: true
format:
  html:
    toc: true
editor_options: 
  chunk_output_type: console
---
# **Instructions:**

For this assignment, you will explore an environmental justice topic of your choosing. You should select a region, community, or environmental issue that matters to you.
* Create two maps that communicate an enviromental justice issue

* Write a brief paragraph explaining what my map communicates.

## I.Setup

```{r}
install.packages("sf")
install.packages("stars")
install.packages("tmap")
```

```{r}
library(tidyverse)
library(sf)
library(here)
library(tmap)
```

## II.Upload data:

```{r}
# reading geodatabase of EJScreen data at the Census Block Group level 
ejscreen <- sf::st_read(here::here("data", "ejscreen","EJSCREEN_2023_BG_StatePct_with_AS_CNMI_GU_VI.gdb")) 

# filtering California
california <- ejscreen %>%
  dplyr::filter(ST_ABBREV == "CA")  

# filter Los Angeles County
#
la_county <- california %>%
  dplyr::filter(CNTY_NAME %in% c("Los Angeles County"))

# find the average values for all variables within counties
california_counties <- aggregate(california, by = list(california$CNTY_NAME), FUN = mean)
```

## III.Create two maps communicating an enviromental justice issue
---

Mapping Number of Low-Income in Los Angeles County, CA:
```{r}
# Static Map
tmap_mode("plot")

# Creates a map of concentration of the % low-income
tm_shape(la_county)





```




## .Analysis
