---
title: "Homework Assignment 1"
subtitle: "Map Making practice"
author: "Amapola Garcia"
date: 2026-10-6
execute: 
  eval: true
format:
  html:
    toc: true
editor_options: 
  chunk_output_type: console
---
# **Instructions:**
For this assignment, you will explore an environmental justice topic of your choosing. You should select a region, community, or environmental issue that matters to you.
---
* Create two maps that communicate an enviromental justice issue
* Write a brief paragraph explaining what my map communicates.

## I.Setup

```{r}
#| eval: false
install.packages("sf")
```

```{r}
#| eval: false
install.packages("sf")
install.packages("stars")
install.packages("tmap")
```

## I.II Upload Libraries
```{r}
library(sf) 
library(stars) 
library(tmap)
library(here)
```

## II.Upload data:
```{r}
# read in geodatabase of EJScreen data at the Census Block Group level
ejscreen <- sf::st_read(here::here("data", "ejscreen","EJSCREEN_2023_BG_StatePct_with_AS_CNMI_GU_VI.gdb")) 

# filter California
california <- ejscreen |> 
  dplyr::filter(ST_ABBREV  == "CA")

# filter to San Mateo County
county <- ejscreen |>
  dplyr::filter(CNTY_NAME %in% c("San Mateo County"))

# find the average values for all variables within counties
california_counties <- aggregate(california, by = list(california$CNTY_NAME), FUN = mean)

```

## III.Create two maps communicating an enviromental justice issue
Mapping the percentage of Low-Income in Los Angeles County, CA:

```{r}
# Static map
tmap_mode("plot")

sm_lwinc <- tm_shape(county) + # defines the spatial object to plot
  tm_polygons(
    size =0.8,
    fill = "LOWINCPCT", #inputs variable from ejscreen data
    fill.scale = tm_scale(values= c("darkorange","lightyellow")),
    fill.legend = tm_legend(title = "% of Low-Income")
    )+
        tm_title("Percentage of Low-Income in San Mateo County")+
        tm_scalebar(position = c("bottom", "left"))+
        tm_compass(position = c("bottom", "right"))
```


```{r}
# Static map
tmap_mode("plot")

tx_air <- tm_shape(county) + 
  tm_polygons(
    size =0.8,
    fill = "RSEI_AIR", #inputs variable from ejscreen
    fill.scale = tm_scale(values= c("-viridis")),
    fill.legend = tm_legend(title = "Levels of Toxic Air Releases")
    )+
        tm_title("Concentration of Toxic Air Releases in San Mateo County")+
        tm_scalebar(position = c("bottom", "left"))+
        tm_compass(position = c("bottom", "right"))
```

## IV. Arrange both maps together
```{r}
tmap_arrange(sm_lwinc, tx_air)
```
## V. Analysis
When thinking about environmental justice issues, two variables from the ejscreen dataset were “% of low income” and "concentration of toxic air releases” within San Mateo County, CA. I wanted to see whether there was a relationship between the percentage of low-income communities and toxic air releases. I asked this because I lived in San Mateo for the last six years before moving to Santa Barbara. San Mateo County is known for being beautiful and green, with lots of nature nearby; however, where I lived, it was predominantly low-income, with industrial sites nearby and close proximity to freeways. Looking at both maps, I see that areas with a higher percentage of low-income people also have a darker shade, indicating higher levels of toxic releases in the same vicinity. I also see that in the second map, where toxic air concentrations are lower, the percentage of low-income people also drops. This suggests structural injustice affecting disadvantaged communities living in certain areas within the county. My maps do have limitations, they show the county as a whole and combine all areas. In my future analysis, I would like to go deeper into the dataset by collecting more information on low-income percentages and showing what types of toxicity are polluting the air. I believe my maps show that in certain regions of the county with higher levels of low-income residents, there are also higher concentrations of toxic air releases.
